---
title: "Unsupervised Learning"
author: "Koji Mizumura"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r prerequisites}

library(tidyverse)
library(magrittr)

```


# Unsupervised learning

## k-means clustering

We have created some two-dimensional data and stored it in a variable called `x` in your workspace. The scatter plot on the right is a visual representation of the data.

In this exercise, your task is to create a k-means model of the x data using 3 clusters, then to look at the structure of the resulting model using the `summary()` function.

```{r}
iris

# Create the k-means model: km.out
km.out <- kmeans(iris[,1:4], centers=3, nstart=20)

# Inspect the result
summary(km.out)
```

## Results of k-means

The `kmeans()` function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the `cluster` component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using `print()` or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model
km.out$cluster

# Print the km.out object
km.out
```

## Visualizing and interpreting results of kmeans()

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard `plot()` function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e. columns) to `plot()` with an extra argument `col = km.out$cluster`, which sets the color of each point in the scatter plot according to its cluster membership.

```{r}
# Scatter plot of x
plot(iris, col = km.out$cluster)
```

## How kmeans() works and practical matters

### Objective
- Explain how k-means algorithm is implemented visually
- Model selection: determining numbers of clusters

### For model selection
- recall k-means has a random component
- best outcome is based on total within cluster sum of squares
  - for each cluster
  - For each observation in the cluster
  - Determine squared distance from the observation to cluster center
  - sum of them altogether
  
We can select `nstart`, running algorithm multiple times helps to find the global minimum total within cluster sum of squares.

```{r eval=FALSE}
# k-means algorithm with 5 centers, run 20 times
kmeans(x, centers =5, nstart=20)
```

### Determine number of clusters

- Trial and error is not the best approach
- One approach is elbow plot. 

## Handling random algorithms 

```{r}
# Set up 2 x 3 plotting grid
par(mfrow = c(2, 3))

# Set seed
set.seed(1)

x <- iris[,1:4]

for(i in 1:6) {
  # Run kmeans() on x with three clusters and one start
  km.out <- kmeans(x, centers = 3, nstart = 1)
  
  # Plot clusters
  plot(x, col = km.out$cluster, 
       main = km.out$tot.withinss)
}
```

## Selecting number of clusters

The k-means algorithm assumes the number of clusters as part of the input. If you know the number of clusters in advance (e.g. due to certain business constraints) this makes setting the number of clusters easy. However, as you saw in the video, if you do not know the number of clusters and need to determine it, you will need to run the algorithm multiple times, each time with a different number of clusters. From this, you can observe how a measure of model quality changes with the number of clusters.

In this exercise, you will run `kmeans()` multiple times to see how model quality changes as the number of clusters changes. Plots displaying this information help to determine the number of clusters and are often referred to as scree plots.

The ideal plot will have an elbow where the quality measure improves more slowly as the number of clusters increases. This indicates that the quality of the model is no longer improving substantially as the model complexity (i.e. number of clusters) increases. In other words, the elbow indicates the number of clusters inherent in the data.


```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# For 1 to 15 cluster centers
for (i in 1:15) {
  km.out <- kmeans(x, centers = i, nstart=20)
  # Save total within sum of squares to wss variable
  wss[i] <- km.out$tot.withinss
}

# Plot total within sum of squares vs. number of clusters
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")
# Set k equal to the number of clusters corresponding to the elbow location
k <- 2
```

## Practical matters: working with real data

Dealing with real data is often more challenging than dealing with synthetic data. Synthetic data helps with learning new concepts and techniques, but the next few exercises will deal with data that is closer to the type of real data you might find in your professional or academic pursuits.

The first challenge with the Pokemon data is that there is no pre-determined number of clusters. You will determine the appropriate number of clusters, keeping in mind that in real data the elbow in the scree plot might be less of a sharp elbow than in synthetic data. Use your judgement on making the determination of the number of clusters.

The second part of this exercise includes plotting the outcomes of the clustering on two dimensions, or features, of the data. These features were chosen somewhat arbitrarily for this exercise. Think about how you would use plotting and clustering to communicate interesting groups of Pokemon to other people.

An additional note: this exercise utilizes the iter.max argument to `kmeans()`. As you've seen, `kmeans()` is an iterative algorithm, repeating over and over until some stopping criterion is reached. The default number of iterations for `kmeans()` is 10, which is not enough for the algorithm to converge and reach its stopping criterion, so we'll set the number of iterations to 50 to overcome this issue. To see what happens when `kmeans()` does not converge, try running the example with a lower number of iterations (e.g. 3). This is another example of what might happen when you encounter real data and use real cases.

```{r}

pokemon <- readr::read_csv("C:/Protected/Data Science/Datacamp/5. Unsupervised learning/Pokemon.csv")

# Initialize total within sum of squares error: wss
wss <- 0
# Look over 1 to 15 possible clusters
for (i in 1:15) {
  # Fit the model: km.out
  km.out <- kmeans(pokemon[,5:12], centers = i, nstart = 20, iter.max = 50)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Select number of clusters
k <- 3

# Build model with k clusters: km.out
km.out <- kmeans(pokemon[,5:12], centers = k, nstart = 20, iter.max = 50)

# View the resulting model
km.out

# Plot of Defense vs. Speed by cluster membership
plot(pokemon[, c("Defense", "Speed")],
     col = km.out$cluster,
     main = paste("k-means clustering of Pokemon with", k, "clusters"),
     xlab = "Defense", ylab = "Speed")
```

# Hierachical clustering

- Number of clusters is not know ahead of time
- Two kinds: bottom-up and top-down, this course bottom-up

```{r}
# calculate similarity as Euclidean distance between observations

dist_matrix <- dist(x)

# retruns hierachical clustering model
hclust(d=dist_matrix)
```

## Hierarchical clustering with results

In this exercise, you will create your first hierarchical clustering model using the `hclust()` function.

We have created some data that has two dimensions and placed it in a variable called `x`. Your task is to create a hierarchical clustering model of `x`. Remember from the video that the first step to hierarchical clustering is determining the similarity between observations, which you will do with the `dist()` function.

You will look at the structure of the resulting model using the `summary()` function.

```{r}
# Create hierarchical clustering model: hclust.out
hclust.out <- hclust(dist(x))

library(magrittr)
x %>% dist() %>% head()

# Inspect the result
summary(hclust.out)
```

## Selecting number of clusters

```{r}
# create hierachical cluster model: hclust.out
hclust.out <- hclust(dist(x))

# inspect the results
summary(hclust.out)
```

### Dendrogram

Tree shaped strcuture used to interpret hierachical clustering models

```{r}
# draw a dendrogram
plot(hclust.out)
abline(h=2, col="red")
```
 
 need to cut the tree to get cluster assigments
 
```{r}
# cut by heright h
cutree(hclust.out, h=2) 

# cut by number of cluster k
cutree(hclust.out, k=6)

```
 
## Cutting the tree

Remember from the video that `cutree()` is the R function that cuts a hierarchical model. The h and k arguments to `cutree()` allow you to cut the tree based on a certain height h or a certain number of clusters k.

In this exercise, you will use `cutree()` to cut the hierarchical model you created earlier based on each of these two criteria.

```{r}
# Cut by height
library(magrittr)
hclust.out %>% cutree(h=7)

# Cut by number of clusters
hclust.out %>% cutree(k=3)
```

## Linkage methods 

In this exercise, you will produce hierarchical clustering models using different linkages and plot the dendrogram for each, observing the overall structure of the trees.

You'll be asked to interpret the results in the next exercise.

```{r}
# Cluster using complete linkage: hclust.complete
hclust.complete <- hclust(dist(x), method = "complete")

# Cluster using average linkage: hclust.average
hclust.average <- hclust(dist(x), method = "average")

# Cluster using single linkage: hclust.single
hclust.single <- hclust(dist(x), method = "single")

# Plot dendrogram of hclust.complete
plot(hclust.complete, main="Complete")


# Plot dendrogram of hclust.average
plot(hclust.average,main="Average")


# Plot dendrogram of hclust.single
plot(hclust.single,
main="Single")

```

## Comparing linkage methods
The models you created in the last exercise— `hclust.complete`, `hclust.average`, and hclust.single—are available in your workspace.

Which linkage(s) produce balanced trees?

## Practical matters: scaling

Recall from the video that clustering real data may require scaling the features if they have different distributions. So far in this chapter, you have been working with synthetic data that did not need scaling.

In this exercise, you will go back to working with "real" data, the pokemon dataset introduced in the first chapter. You will observe the distribution (mean and standard deviation) of each feature, scale the data accordingly, then produce a hierarchical clustering model using the complete linkage method.

```{r eval=FALSE}
# View column means
library(magrittr)
pokemon %>% colMeans()

# View column standard deviations
apply(pokemon, sd, MARGIN=2)

# Scale the data
pokemon.scaled <- scale(pokemon)

# Create hierarchical clustering model: hclust.pokemon
hclust.pokemon = hclust(dist(pokemon.scaled), method="complete")
```

## Comparing kmeans() and hclust()
Comparing k-means and hierarchical clustering, you'll see the two methods produce different cluster memberships. This is because the two algorithms make different assumptions about how the data is generated. In a more advanced course, we could choose to use one model over another based on the quality of the models' assumptions, but for now, it's enough to observe that they are different.

```{r eval=FALSE}
# Apply cutree() to hclust.pokemon: cut.pokemon
cut.pokemon <- cutree(hclust.pokemon, k=3)

# Compare methods
table(km.pokemon$cluster, cut.pokemon)
```

# Dimensionality reduction by PCA
## Introduction to PCA

- Two methods of clustering - finding groups of homogeneous items
- Next up, dimensionality reduction

Dimensionality reduction is 

- A popular method is PCA
- Three goals when finding lower dimensional representation of features
  - Find linear combination of variables to create principla components
  - Maintain most variance in the data
  - Principal componnets are uncorrelated (i.e., orthogonal to each other)
  
```{r}
pr.iris <- prcomp(
  x = iris[-5],
  scale = FALSE,
  center = TRUE
)

summary(pr.iris)

pr.iris$x %>% 
  plot()
```

## PCA using prcomp()
In this exercise, you will create your first PCA model and observe the diagnostic results.

We have loaded the Pokemon data from earlier, which has four dimensions, and placed it in a variable called pokemon. Your task is to create a PCA model of the data, then to inspect the resulting model using the `summary()` function.

```{r}
# Perform scaled PCA: pr.out
pokemon <- iris[-5]

pr.out <- prcomp(pokemon, scale=TRUE)

# Inspect model output
summary(pr.out)
```

## Results of PCA
This exercise will check your understanding of the `summary()` of a PCA model. Your model from the last exercise, pr.out, and the pokemon dataset are still available in your workspace.

What is the minimum number of principal components that are required to describe at least 75% of the cumulative variance in this dataset?

## Additional results of PCA
PCA models in R produce additional diagnostic and output components:

- `center`: the column means used to center to the data, or FALSE if the data weren't centered
- `scale`: the column standard deviations used to scale the data, or FALSE if the data weren't scaled
- `rotation`: the directions of the principal component vectors in terms of the original features/variables. This information allows you to define new data in terms of the original principal components
- `x`: the value of each observation in the original dataset projected to the principal components

You can access these the same as other model components. For example, use pr.out$rotation to access the rotation component.

Which of the following statements is not correct regarding the pr.out model fit on the pokemon data?

The directions of the principal component vectors are presented in a table with the same dimensions as the original data

## Variance explained

The second common plot type for understanding PCA models is a scree plot. A scree plot shows the variance explained as the number of principal components increases. Sometimes the cumulative variance explained is plotted as well.

In this and the next exercise, you will prepare data from the pr.out model you created at the beginning of the chapter for use in a scree plot. Preparing the data for plotting is required because there is not a built-in function in R to create this type of plot.

```{r}
# Variability of each principal component: pr.var
pr.var <- pr.out$sdev^2

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)
```

## Visualize variance explained

Now you will create a scree plot showing the proportion of variance explained by each principal component, as well as the cumulative proportion of variance explained.

Recall from the video that these plots can help to determine the number of principal components to retain. One way to determine the number of principal components to retain is by looking for an elbow in the scree plot showing that as the number of principal components increases, the rate at which variance is explained decreases substantially. In the absence of a clear elbow, you can use the scree plot as a guide for setting a threshold.

```{r}
# Plot variance explained for each principal component
library(magrittr)
pve
plot(pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")
```

## Practical issues with PCA

- Scaling the data:
- Missing values:
  - Drop observations with missing values
  - Impute / estimate missing values
- categorical data:
  - Do not use categorical data features
  - Encode categorical features as numbers
  
  
```{r}
data(mtcars)
head(mtcars)
```
  
```{r}
# Means and standard deviation vary a lot
round(colMeans(mtcars),2)
round(apply(mtcars, 2,sd),2)
```

Scaling and PCA in R

```{r eval=FALSE}
prcomp(x, center = TRUE, scale = FALSE)
```

## Practical issues: scaling

You saw in the video that scaling your data before doing PCA changes the results of the PCA modeling. Here, you will perform PCA with and without scaling, then visualize the results using biplots.

Sometimes scaling is appropriate when the variances of the variables are substantially different. This is commonly the case when variables have different units of measurement, for example, degrees Fahrenheit (temperature) and miles (distance). Making the decision to use scaling is an important step in performing a principal component analysis.

```{r}
# Mean of each variable
colMeans(pokemon)

# Standard deviation of each variable
apply(pokemon, 2, sd)

# PCA model with scaling: pr.with.scaling
pr.with.scaling <- prcomp(pokemon, scale=TRUE)

# PCA model without scaling: pr.without.scaling
pr.without.scaling <- prcomp(pokemon, scale=FALSE)

# Create biplots of both for comparison
biplot(pr.with.scaling)
biplot(pr.without.scaling)

```

# Putting it all together with a case study
## Introduction

- Download data and prepare data for modeling
- Exploratory data analysis (#observation, #features, etc.)
- Perform PCa nad interpret results
- Complete two types of clustering
- Inderstand and compare two types
- Combine PCA and clustering

## Preparing the data
```{r}
url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1903/datasets/WisconsinCancer.csv"

# Download the data: wisc.df
wisc.df <- read.csv(url)

# Convert the features of the data: wisc.data
wisc.data = as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

# Create diagnosis vector
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```


## EDA

The first step of any data analysis, unsupervised or supervised, is to familiarize yourself with the data.

The variables you created before, wisc.data and diagnosis, are still available in your workspace. Explore the data to answer the following questions:

1. How many observations are in this dataset?
2. How many variables/features in the data are suffixed with _mean?
3. How many of the observations have a malignant diagnosis?
```{r}
wisc.df
wisc.df %>% dim()
```

## Peform PCA

The next step in your analysis is to perform PCA on wisc.data.

You saw in the last chapter that it's important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data:

1. The input variables use different units of measurement.
2. The input variables have significantly different variances.

```{r}
# Check column means and standard deviations
dim(wisc.data)
colMeans(wisc.data)
apply(wisc.data, 2, sd)

# Execute PCA, scaling if appropriate: wisc.pr
wisc.pr = prcomp(wisc.data, scale=TRUE)

# Look at summary of results
summary(wisc.pr)
```

## Interpreting PCA results

Now you'll use some visualizations to better understand your PCA model. You were introduced to one of these visualizations, the biplot, in an earlier chapter.

You'll run into some common challenges with using biplots on real-world data containing a non-trivial number of observations and variables, then you'll look at some alternative visualizations. You are encouraged to experiment with additional visualizations before moving on to the next exercise.

```{r}
# Create a biplot of wisc.pr
biplot(wisc.pr)
wisc.pr$x

# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[, c(1, 2)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC2")

# Repeat for components 1 and 3
plot(wisc.pr$x[,c(1,3)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")

# Do additional data exploration of your choosing below (optional)

```

## Variance explained

n this exercise, you will produce scree plots showing the proportion of variance explained as the number of principal components increases. The data from PCA must be prepared for these plots, as there is not a built-in function in R to create them directly from the PCA model.

As you look at these plots, ask yourself if there's an elbow in the amount of variance explained that might lead you to pick a natural number of principal components. If an obvious elbow does not exist, as is typical in real-world datasets, consider how else you might determine the number of principal components to retain based on the scree plot.

```{r}
# Set up 1 x 2 plotting grid
par(mfrow = c(1, 2))

# Calculate variability of each component
summary(wisc.pr)
pr.var <- wisc.pr$sdev^2
pr.var
# Variance explained by each principal component: pve
pve <- pr.var/sum(pr.var) 

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")
```

## Communicating PCA results

This exercise will check your understanding of the PCA results, in particular the loadings and variance explained. The loadings, represented as vectors, explain the mapping from the original features to the principal components. The principal components are naturally ordered from the most variance explained to the least variance explained.

The variables you created before—wisc.data, diagnosis, wisc.pr, and pve—are still available.

For the first principal component, what is the component of the loading vector for the feature concave.points_mean? What is the minimum number of principal components required to explain 80% of the variance of the data?

## PCA review and next steps

- Complete hierchacical clustering
- Complete k-means clustering
- Combine PCA and clustering
- contrast results of hierchacial clustering with diagnosis
- compare hierachical and k-means clustering reuslts
- PCA as a pre-procesing step for clustering

```{r}
# pair-wise distance between observation - Euclidiean dsitance
s <- dist(x)
hclust(x)

# one observation per row, one feature per column
# k-means has arandom component
# run algorithm multiple times to improve odds of the best model

```






