---
title: "ML in Tidyverse"
author: "Koji Mizumura"
date: "2019-3-12 - `r Sys.Date()`"
output: 
  html_document:
    number_sections: yes
    section_divs: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  fig.height = 4.5,
  error = FALSE,
  message = FALSE, 
  warning = FALSE, 
  collapse = TRUE,
  cache = TRUE
)
```

# Foundations of "tidy" ML

```{r pre-requisites, include=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(tidymodels)
```


## Nesting the data

List column workflow is as follows.

1. Make a list column: `nest()`
2. Work with __list columns__: `map()`
3. Simplify the list columns `unnest()` and `map_*`

In this course, you will work with a collection of economic and social indicators for 77 countries over a period of 52 years. This data is stored in the `gapminder` dataframe.

In this exercise, you will transform your `gapminder` data into a nested dataframe by using the first tool needed to build the foundation of tidy machine learning skills: `nest()`.

Note: This is a more granular version than the dataset available from the gapminder package. This version is available in the `dslabs` package.

```{r warning=FALSE}
library(gapminder)

# Explore gapminder
head(gapminder)

# Prepare the nested dataframe gap_nested
library(tidyverse)
gap_nested <- gapminder %>% 
  group_by(country) %>% 
  nest()

# Explore gap_nested
head(gap_nested)
```


## Unnesting your data

As you've seen in the previous exercise, a nested dataframe is simply a way to shape your data. Essentially taking the `group_by()` windows and packaging them in corresponding rows.

In the same way you can use the `nest()` function to break your data into nested chunks, you can use the `unnest()` function to expand the dataframes that are nested in these chunks.
```{r}
# Create the unnested dataframe called gap_unnnested
gap_unnested <- gap_nested %>% 
  unnest()
  
# Confirm that your data was not modified  
identical(gapminder, gap_unnested)
```

## Explore a nested cell

In the first exercise, you successfully created a nested dataframe `gap_nested`. The `data` column contains tibbles for each country. In this exercise, you will explore one of these nested chunks.

```{r}
# Extract the data of Algeria
algeria_df <- gap_nested$data[[1]]
algeria_df %>% colnames()
# Calculate the minimum of the population vector
min(algeria_df$population)

# Calculate the maximum of the population vector
max(algeria_df$population)

# Calculate the mean of the population vector
mean(algeria_df$population)
```

## Map() function

$$
map(.x = , .f = )
$$

- .x = `[vector]` or `[[list]]`
- .f = `mean` or `~mean(.x)` 

## Mapping your data

In combination with `mutate()`, you can use `map()` to append the results of your calculation to a dataframe. Since the `map()` function always returns a vector of lists you must use `unnest()` to extract this information into a numeric vector.

Here you will explore this functionality by calculating the mean population of each country in the gapminder dataset.

```{r}
# Calculate the mean population for each country
gap_nested
pop_nested <- gap_nested %>%
  mutate(mean_pop = map(data, ~mean(.x$population)))

# Take a look at pop_nested
head(pop_nested)

# Extract the mean_pop value by using unnest
pop_mean <- pop_nested %>% 
  unnest(mean_pop)

# Take a look at pop_mean
head(pop_mean)
```

```{r}
# Calculate mean population and store result as a double
pop_mean <- gap_nested %>%
  mutate(mean_pop = map_dbl(data, ~mean(.x$population)))

# Take a look at pop_mean
head(pop_mean)
```

## Mapping many models

The `gap_nested` dataframe available in your workspace contains the gapminder dataset nested by country.

You will use this data to build a linear model for each country to predict life expectancy using the year feature.

Note: The term feature is synonymous with the terms variable or predictor. It refers to an attribute of your data that can be used to build a machine learning model.

```{r}
# Build a linear model for each country
gap_models <- gap_nested %>%
    mutate(model = map(data, ~lm(formula = lifeExp~year, data = .x)))
    
# Extract the model for Algeria    
algeria_model <- gap_models$model[[1]]

# View the summary for the Algeria model
summary(algeria_model)
```

## tidying models with broom

To Work with list columns, we use `broom`, `Metrics`, `rsample` package s etc.

- `tidy()`: returns the statistical findings of the model (such as coefficients)
- `glance()`: returns a concise one-row summary of the model
- `augment()`: adds prediction columns to the data being modeled

```{r}
algeria_model %>% broom::tidy()
algeria_model %>% glance()
algeria_model %>% augment()
```

## Extracting model statistics tidily
In this exercise, you will use the `tidy()` and `glance()` functions to extract information from algeria_model in a tidy manner.

For a linear model, `tidy()` extracts the model coefficients while `glance()` returns the model statistics such as the $R^2$

```{r}
library(broom)
library(magrittr)

# Extract the coefficients of the algeria_model as a dataframe
broom::tidy(algeria_model)
  
# Extract the statistics of the algeria_model as a dataframe
glance(algeria_model)
```

## Augmenting your data
From the results of `glance()`, you learned that using the available features the linear model fits well with an adjusted $R^2$ of $0.99$. The `augment()` function can help you explore this fit by appending the predictions to the original data.

Here you will leverage this to compare the predicted values of `life_expectancy` with the original ones based on the `year` feature.

```{r}
# Build the augmented dataframe
algeria_fitted <- augment(algeria_model)

# Compare the predicted values with the actual values of life expectancy
algeria_fitted %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = lifeExp)) + 
  geom_line(aes(y = .fitted), color = "red")
```

# Multiple models with broom
## Exploring coefficients across models

Tidy up the coefficients of your models
In this exercise you will leverage the list column workflow along with the `tidy()` function from broom to extract and explore the coefficients for the 77 models you built.

Remember the `gap_models` dataframe contains a model predicting life expectancy by year for 77 countries.

```{r}
# Extract the coefficient statistics of each model into nested dataframes
model_coef_nested <- gap_models %>% 
    mutate(coef = map(model, ~broom::tidy(.x)))
    
# Simplify the coef dataframes for each model    
model_coef <- model_coef_nested %>%
    unnest(coef)

# Plot a histogram of the coefficient estimates for year         
model_coef %>% 
  filter(term == "year") %>% 
  ggplot(aes(x =estimate)) +
  geom_histogram()
```

## What can we learn about these 77 countries?
Explore the `model_coef` dataframe you just created to answer the following question:

Which of the following conclusions can we make from the coefficients of our models?

## Glance at the fit of your model

In this exercise you will use `glance()` to calculate how well the linear models fit the data for each country.

```{r}
# Extract the fit statistics of each model into dataframes
model_perf_nested <- gap_models %>% 
    mutate(fit = map(model, ~glance(.x)))

# Simplify the fit dataframes for each model    
model_perf <- model_perf_nested %>% 
    unnest(fit)
    
# Look at the first six rows of model_perf
head(model_perf)
```

## Best and worst fitting models

```{r}
# Plot a histogram of rsquared for the 77 models    
model_perf

model_perf %>% 
  ggplot(aes(x = r.squared)) + 
  geom_histogram()  
  
# Extract the 4 best fitting models
best_fit <- model_perf %>% 
  top_n(n = 4, wt = r.squared)

# Extract the 4 models with the worst fit
worst_fit <- model_perf %>% 
  top_n(n = 4, wt = -r.squared)
```

## Augment the fitted values of each model

In this exercise you will prepare your four best and worst fitting models for further exploration by augmenting your model data with `augment()`.

```{r}
best_augmented <- best_fit %>% 
  # Build the augmented dataframe for each country model
  mutate(augmented = map(model, ~augment(.x))) %>% 
  # Expand the augmented dataframes
  unnest(augmented)

worst_augmented <- worst_fit %>% 
  # Build the augmented dataframe for each country model
  mutate(augmented = map(model, ~augment(.x))) %>% 
  # Expand the augmented dataframes
  unnest(augmented)
```

## Explore your best and worst fitting models
Let's explore your four best and worst fitting models by comparing the fitted lines with the actual values.

```{r}
# Compare the predicted values with the actual values of life expectancy 
# for the top 4 best fitting models
best_augmented %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = lifeExp)) + 
  geom_line(aes(y = .fitted), color = "red") +
  facet_wrap(~country, scales = "free_y")

# Compare the predicted values with the actual values of life expectancy 
# for the top 4 worst fitting models
worst_augmented %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = lifeExp)) + 
  geom_line(aes(y = .fitted), color = "red") +
  facet_wrap(~country, scales = "free_y")



```

To improve model fits...

## Build better models

Earlier you built a collection of simple models to fit life expectancy using the year feature. Your previous analysis showed that some of these models didn't fit very well.

In this exercise you will build multiple regression models for each country using all available features. You may be interested in comparing the performance of the four worst fitting models so their adjusted $R^2$ are provided below:

```{r eval=FALSE}
# Build a linear model for each country using all features

gap_fullmodel <- gap_nested %>% 
  mutate(model = map(data, ~lm(formula = lifeExp ~ . , data = .x)))

fullmodel_perf <- gap_fullmodel %>% 
  # Extract the fit statistics of each model into dataframes
  mutate(fit = map(model, ~glance(.x))) %>% 
  # Simplify the fit dataframes for each model
  unnest(fit)
  
# View the performance for the four countries with the worst fitting 
# four simple models you looked at before
fullmodel_perf %>% 
  filter(country %in% worst_fit$country) %>% 
  select(country, adj.r.squared)
```

# Build Tune, Evaluate Regression Models
## The test-train split

In a disciplined machine learning workflow it is crucial to withhold a portion of your data (__testing data__) from any decision-making process. This allows you to independently assess the performance of your model when it is finalized. The remaining data, the __training data__, is used to build and select the best model.

In this exercise, you will use the `rsample` package to split your data to perform the initial train-test split of your `gapminder` data.

Note: Since this is a random split of the data it is good practice to set a seed before splitting it.

```{r}
set.seed(42)

# Prepare the initial split object
gap_split <- initial_split(gapminder, prop = 0.75)

# Extract the training dataframe
training_data <- training(gap_split)

# Extract the testing dataframe
testing_data <- testing(gap_split)

# Calculate the dimensions of both training_data and testing_data
dim(training_data)
dim(testing_data)
```

## Cross-validation dataframes
Now that you have withheld a portion of your data as __testing data__, you can use the remaining portion to find the best performing model.

In this exercise, you will split the training data into a series of 5 train-validate sets using the `vfold_cv()` function from the `rsample` package.

```{r}
set.seed(42)

# Prepare the dataframe containing the cross validation partitions
cv_split <- vfold_cv(training_data, v = 5)
cv_split

cv_data <- cv_split %>% 
  mutate(
    # Extract the train dataframe for each split
    train = map(splits, ~training(.x)), 
    # Extract the validate dataframe for each split
    validate = map(splits, ~testing(.x))
  )

# Use head() to preview cv_data
head(cv_data)
```

## Measuring cross-validation performance

- MAE: How much on average the model's prediction differ from actual observations.

Three steps to calculate MAE: 
1) Build cross-validated models
2) Predict using trained models by `map2`
3) Compute MAE by `map2_dbl`

## Build cross-validated models
In this exercise, you will build a linear model predicting life_expectancy using all available features. You will do this for the train data of each cross-validation fold.

```{r}
# Build a model using the train data for each fold of the cross validation
cv_data

cv_models_lm <- cv_data %>% 
  mutate(model = map(train, ~lm(formula = lifeExp ~., data = .x)))
```

## Preparing for evaluation
In order to measure the validate performance of your models you need compare the predicted values of `life_expectancy` for the observations from validate set to the actual values recorded. Here you will prepare both of these vectors for each partition.

```{r}
cv_prep_lm <- cv_models_lm %>% 
  mutate(
    # Extract the recorded life expectancy for the records in the validate dataframes
    validate_actual = map(validate, ~.x$lifeExp),
    # Predict life expectancy for each validate set using its corresponding model
    validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y))
  )

cv_prep_lm
```

## Evaluate model performance
Now that you have both the actual and predicted values of each fold you can compare them to measure performance.

For this regression model, you will measure the __Mean Absolute Error (MAE)__ between these two vectors. This value tells you the average difference between the actual and predicted values.

```{r}
library(Metrics)
# Calculate the mean absolute error for each validate fold       
cv_eval_lm <- cv_prep_lm %>% 
  mutate(validate_mae = map2_dbl(validate_actual, validate_predicted, ~mae(actual = .x, predicted = .y)))

# Print the validate_mae column
cv_eval_lm$validate_mae

# Calculate the mean of validate_mae column
mean(cv_eval_lm$validate_mae)
```

## Building and tuning a random forest model

As another model, the random forest is beneficial for 1) handling non-linear relationships, 2) handling interactions.

Model
```{r eval=FALSE}
rf_model <- ranger(
  formula = XX,
  data = XX,
  seed = XX
)
```

Prediction
```{r eval=FALSE}
prediction <- predict(
  rf_model,
  new_data
)$predictions
```

For random forest model, hyper parameters are

- `mtry`: default is $\sqrt{number of features}$, and the range is $1:number of features$
- `num.trees`: default is $500$, and the range is $1:\infty$.

## Build a random forest model by `ranger`

Here you will use the same cross-validation data to build (using `train`) and evaluate (using `validate`) random forests for each partition. Since you are using the same cross-validation partitions as your regression models, you are able to directly compare the performance of the two models.

Note: We will limit our random forests to contain $100$ trees to ensure they finish fitting in a reasonable time. The _default number of trees_ for `ranger()` is $500$.

```{r}
library(ranger)

# Build a random forest model for each fold

cv_models_rf <- cv_data %>% 
  mutate(model = map(train, ~ranger(formula = lifeExp ~., data = .x,
                                    num.trees = 100, seed = 42)))

# Generate predictions using the random forest model
cv_data
cv_models_rf

cv_prep_rf <- cv_models_rf %>% 
  mutate(
    validate_actual = map(validate, ~.x$lifeExp),
    validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y)$predictions))
cv_prep_rf
```

## Evaluate a random forest model
Similar to the linear regression model, you will use the MAE metric to evaluate the performance of the random forest model.

```{r}
library(ranger)

# Calculate validate MAE for each fold
cv_eval_rf <- cv_prep_rf %>% 
  mutate(validate_mae = map2_dbl(validate_actual, validate_predicted, ~mae(actual = .x, predicted = .y)))

# Print the validate_mae column
cv_eval_rf$validate_mae

# Calculate the mean of validate_mae column
mean(cv_eval_rf$validate_mae)
```

## Fine tune your model
Wow! That was a significant improvement over a regression model. Now let's see if you can further improve this performance by fine tuning your random forest models. To do this you will vary the `mtry` parameter when building your random forest models on your `train` data.

The default value of `mtry` for ranger is the rounded down square root of the total number of features (6). This results in a value of 2.

```{r}
# Prepare for tuning your cross validation folds by varying mtry
cv_tune <- cv_data %>% 
  mutate(validate_actual = map(validate, ~.x$lifeExp)) %>% 
  crossing(mtry = 2:5) 

# Build a model for each fold & mtry combination
library(ranger)

cv_model_tunerf <- cv_tune %>% 
  mutate(model = map2(.x = train, .y = mtry, ~ranger(formula = lifeExp~., 
                                           data = .x, mtry = .y, 
                                           num.trees = 100, seed = 42)))
```

## The best performing parameter

You've now built models where you've varied the random forest-specific hyperparameter `mtry` in the hopes of improving your model further. Now you will measure the performance of each `mtry` value across the 5 cross validation partitions to see if you can improve the model.

Remember that the validate MAE you calculated two exercises ago of `0.795` was for the default mtry value of 2.

```{r eval=FALSE}
# Generate validate predictions for each model
cv_model_tunerf

cv_prep_tunerf <- cv_model_tunerf %>% 
  mutate(validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y)$predictions))

# Calculate validate MAE for each fold and mtry combination
cv_eval_tunerf <- cv_prep_tunerf %>% 
  mutate(validate_mae = map2_dbl(.x = validate_predicted, .y = validate_actual, ~mae(actual = .x, predicted = .y)))

# Calculate the mean validate_mae for each mtry used  
cv_eval_tunerf %>% 
  group_by(mtry) %>% 
  summarise(mean_mae = mean(validate_mae))
```

## 

The test portion was intentionally held out in order to evaluate the final model with an independent set of data. The train portion of data is further split into train and validate using cross validation for the purpose of model selection.

Each train portion was build a model and the held out validate protion was used to evaluate it. Resulting in measures of validation performance for each cross validation fold for each model and hyperparameter.

Aggreagting the validation performance for each model allowed us to compare multiple models as well as their respective hyperparameters to select model hyper-parameter combination with the best overall performance.

```{r}
# Build the model using all training data and the best performing parameter
best_model <- ranger(formula = lifeExp~., data = training_data,
                     mtry = 4, num.trees = 100, seed = 42)

# Prepare the test_actual vector
test_actual <- testing_data$lifeExp

# Predict life_expectancy for the testing_data
test_predicted <- predict(best_model, testing_data)$predictions

# Calculate the test MAE
mae(test_actual, test_predicted)
```

# Build, Tune & Evaluate classification models

## Prepare train-test-validate parts

In this exercise, you will leverage the tools that you have learned thus far to build a classification model to predict employee attrition.

You will work with the `attrition` dataset, which contains 30 features about employees which you will use to predict if they have left the company.

You will first prepare the training & testing data sets, then you will further split the training data using cross-validation so that you can search for the best performing 

```{r}
set.seed(42)
# Prepare the initial split object
data_split <- initial_split(attrition, prop = 0.75)

# Extract the training dataframe
training_data <- training(data_split)

# Extract the testing dataframe
testing_data <- testing(data_split)


set.seed(42)
cv_split <- vfold_cv(training_data, v = 5)

cv_split

cv_data <- cv_split %>% 
  mutate(
    # Extract the train dataframe for each split
    train = map(splits, ~training(.x)),
    # Extract the validate dataframe for each split
    validate = map(splits, ~testing(.x))
  )

cv_data
```

## Build cross-validated models

In this exercise, you will build logistic regression models for each fold in your cross-validation.

You will build this using the `glm()` function and by setting the family argument to `"binomial"`.

```{r}
# Build a model using the train data for each fold of the cross validation

cv_models_lr <- cv_data %>% 
  mutate(model = map(train, ~glm(formula = Attrition~., 
                               data = .x, family = "binomial")))

cv_models_lr
```

## Evaluating classification models

The ingredients needed to measure performance are the same as before. 

- Actual classes of your observations: actual `Attrition` classes
- Predicted classes: predicted `attrition` classes
- A metric to compare 1) and 2).

As a metric, `accuracy`, `precision` and `recall` can be used.

- `accuracy`: how well your model predicted both the TRUE and FALSE classes
- `precision`: appropriate when you want to minimize how often the model incorrectly predicts an observation to be in the positive class
- `recall`: This metrics compares the number of observations the model has correctly identified as TRUE to the total number of TRUE observations.In other words, it measures the rate at which the model can capture the TRUE class. This metric would be appropriate when building a model that would capture as many risky employees as possible you should consider this metric.

## Predicitons of a single model

To calculate the performance of a classification model you need to compare the actual values of `Attrition` to those predicted by the model. When calculating metrics for binary classification tasks (such as precision and recall), the actual and predicted vectors must be converted to __binary__ values.

In this exercise, you will learn how to prepare these vectors using the model and validate dataframes from the first cross-validation fold as an example.

```{r}

# Extract the first model and validate 
model <- cv_models_lr$model[[1]]
validate <- cv_models_lr$validate[[1]]

# Prepare binary vector of actual Attrition values in validate
validate_actual <- validate$Attrition == "Yes"

# Predict the probabilities for the observations in validate
validate_prob <- predict(model, validate, type = "response")

# Prepare binary vector of predicted Attrition values for validate
validate_predicted <- validate_prob > 0.5
```

## cv_models_lr

Now that you have the binary vectors for the actual and predicted values of the model, you can calculate many commonly used binary classification metrics. In this exercise you will focus on:

- __accuracy__: rate of correctly predicted values relative to all predictions.
- __precision__: portion of predictions that the model correctly predicted as TRUE.
- __recall__: portion of actual TRUE values that the model correctly recovered.

```{r}
library(Metrics)

# Compare the actual & predicted performance visually using a table
table(validate_actual, validate_predicted)

# Calculate the accuracy
accuracy(validate_actual, validate_predicted)

# Calculate the precision
precision(validate_actual, validate_predicted)

# Calculate the recall
recall(validate_actual, validate_predicted)
```

## Prepare for cross-validated performance

Now that you know how to calculate the performance metrics for a single model, you are now ready to expand this for all the folds in the cross-validation dataframe.

```{r}
cv_models_lr$validate[[1]]


cv_prep_lr <- cv_models_lr %>% 
  mutate(
    # Prepare binary vector of actual Attrition values in validate
    validate_actual = map(validate, ~.x$Attrition == "Yes"),
    # Prepare binary vector of predicted Attrition values for validate
    validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y, type = "response") > 0.5)
  )
```

## Calculate cross-validated performance

It is crucial to optimize models using a carefully selected metric aimed at achieving the goal of the model.

Imagine that in this case you want to use this model to identify employees that are predicted to leave the company. Ideally, you want a model that can capture as many of the ready-to-leave employees as possible so that you can intervene. The corresponding metric that captures this is the recall metric. As such, you will exclusively use recall to optimize and select your models.

```{r}
# Calculate the validate recall for each cross validation fold

cv_prep_lr %>% colnames()

cv_perf_recall <- cv_prep_lr %>% 
  mutate(validate_recall = map2_dbl(validate_actual, validate_predicted, 
                                    ~recall(actual = .x, predicted = .y)))

# Print the validate_recall column
cv_perf_recall$validate_recall

# Calculate the average of the validate_recall column
mean(cv_perf_recall$validate_recall)
```

## Classification with random forests

`ranger()` for classification. Tuning and building process is same as before. Since there are 30 features in the Attrition dataset, this value can go as high as 30.For now, we will try a few `mtry` values.

## Tune random forest models

Now that you have a working logistic regression model you will prepare a random forest model to compare it with.

```{r}
library(ranger)

# Prepare for tuning your cross validation folds by varying mtry
cv_tune <- cv_data %>%
  crossing(mtry = c(2,4,8,16)) 

cv_tune

# Build a cross validation model for each fold & mtry combination
cv_models_rf <- cv_tune %>% 
  mutate(model = map2(train, mtry, ~ranger(formula = Attrition~., 
                                           data = .x, mtry = .y,
                                           num.trees = 100, seed = 42)))
```

## Random forest performance
It is now time to see whether the random forests models you built in the previous exercise are able to outperform the logistic regression model.

Remember that the validate recall for the logistic regression model was 0.43.

```{r }
cv_prep_rf <- cv_models_rf %>% 
  mutate(
    # Prepare binary vector of actual Attrition values in validate
    validate_actual = map(validate, ~.x$Attrition == "Yes"),
    # Prepare binary vector of predicted Attrition values for validate
    validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y, type = "response")$predictions == "Yes")
  )

cv_prep_rf

# Calculate the validate recall for each cross validation fold
cv_perf_recall <- cv_prep_rf %>% 
  mutate(recall = map2_dbl(.x = validate_actual, .y = validate_predicted, ~recall(actual = .x, predicted = .y)))

# Calculate the mean recall for each mtry used  
cv_perf_recall %>% 
  group_by(mtry) %>% 
  summarise(mean_recall = mean(recall))
```

## Build final classification model
Comparing the recall performance between the logistic regression model (0.4) and the best performing random forest model (0.2), you've learned that the model with the best performance is the logistic regression model. In this exercise, you will build the logistic regression model using all of the train data and you will prepare the necessary vectors for evaluating this model's test performance.

```{r}
# Build the logistic regression model using all training data
best_model <- glm(formula = Attrition~., 
                  data = training_data, family = "binomial")

# Prepare binary vector of actual Attrition values for testing_data
test_actual <- testing_data$Attrition == "Yes"

# Prepare binary vector of predicted Attrition values for testing_data
test_predicted <- predict(best_model, testing_data, type = "response") > 0.5
```

## Measure final model performance
Now its time to calculate the test performance of your final model (logistic regression). Here you will use the held out testing data to characterize the performance you would expect from this model when it is applied to new data.

```{r}
# Compare the actual & predicted performance visually using a table
table(test_actual, test_predicted)

# Calculate the test accuracy
accuracy(test_actual, test_predicted)

# Calculate the test precision
precision(test_actual, test_predicted)

# Calculate the test recall
recall(test_actual, test_predicted)
```

